\vspace{-1.5em} \bf A general and efficient representation of ancestral recombination graphs

### 2023-10-26 paper.tex
Edits from Yan and JEre

% This is nice to have, but probably won't work in the final version
% without section numbers. Let's deal with that when we get there I guess.
We begin in Section~\ref{sec-gARG} by providing a precise formal definition
of a gARG, illustrated by an example ARG embedded in pedigree.
We then provide a similar definition of the classical eARG
approach in Section~\ref{sec-eARG}, and consider its limitations
in the context of contemporary datasets.
We then discuss the import concept of ``ancestral material''
in Section~\ref{sec-ancestral-material}, and how this relates
to the process of converting an eARG to a gARG.
Using this new definition for the remainder of the paper,
we continue
in Section\ref{sec-ARG-and-local-trees} by considering the relationship
between an ARG and its local trees and show (contrary to the
% I want to put a (suitably encoded) here somewhere, but the sentence is
% very long already.
prevailing view) that the sequence of local trees contains
precisely as much information as the corresponding ARG.
The gARG view also opens a rich new set of details about ARGs,
including the ideas of locally unary nodes
(Section~\ref{sec-locally-unary-edges}),
the levels of detail that can be represented in an ARG
(Section~\ref{sec-ARG-simplification}),
and (relatedly) the degrees of precision about recombination
that can be stored (Section~\ref{sec-precision}).
These ideas have important practical considerations,
which we illustrate by examining the qualitative properties of the
ARGs inferred by four recent methods for a classical benchmark
dataset in Section~\ref{sec-example-inferred-args}.
We then discuss how the gARG framework can be
efficiently implementated in Section~\ref{sec-efficiency},
and finish with an assessment of the key challenges facing
the field in the Discussion.
Finally, the literature on ARGs and associated aspects is large and
confusing, and we cover some important aspects
in appendices, including
a brief history of ancestral graphs (Appendix~\ref{sec-arg-history}),
a description of the Big and Little ARG stochastic processes
 (Appendix~\ref{sec-big-and-little-arg}),
a survey of ARG inference methods (Appendix~\ref{sec-survey-arg-infer}),
and a discussion of ARGs at an individual vs cell
lineage level (Appendix~\ref{sec-cell-lineages-and-args}).

>>>>

% This is nice to have, but probably won't work in the final version
% without section numbers. Let's deal with that when we get there I guess.
We begin in Section~\ref{sec-gARG} by providing a precise formal definition
of a gARG, illustrated by an example ARG embedded in pedigree.
We then provide a similar definition of the classical eARG
approach in Section~\ref{sec-eARG}, and consider its limitations
in the context of contemporary datasets.
We then discuss the import concept of ``ancestral material''
in Section~\ref{sec-ancestral-material}, and how this relates
to the process of converting an eARG to a gARG.
Using this new definition for the remainder of the paper,
we continue
in Section~\ref{sec-ARG-and-local-trees} by considering the relationship
between an ARG and its local trees. Contrary to the
prevailing view, we show that a (suitably encoded) sequence of local trees contains
precisely as much information as the corresponding ARG.
The gARG view also opens a rich new set of details about ARGs,
including the ideas of locally unary nodes
(Section~\ref{sec-locally-unary-edges}),
the levels of detail that can be represented in an ARG
(Section~\ref{sec-ARG-simplification}),
and (relatedly) the degrees of precision about recombination
that can be stored (Section~\ref{sec-precision}).
These ideas have important practical considerations,
which we illustrate by examining the qualitative properties of the
ARGs inferred by four recent methods for a classical benchmark
dataset in Section~\ref{sec-example-inferred-args}.
We then discuss how the gARG framework can be
efficiently implementated in Section~\ref{sec-efficiency},
and finish with an assessment of the key challenges facing
the field in the Discussion.
Finally, the literature on ARGs and associated aspects is large and
confusing, and we cover some important aspects
in appendices, including
a brief history of ancestral graphs (Appendix~\ref{sec-arg-history}),
a description of the Big and Little ARG stochastic processes
 (Appendix~\ref{sec-big-and-little-arg}),
a survey of ARG inference methods (Appendix~\ref{sec-survey-arg-infer}),
and a discussion of ARGs at an individual vs cell
lineage level (Appendix~\ref{sec-cell-lineages-and-args}).
